Spatial variation in allometric growth of invasive lionfish has management implications

### 2018-07-13 docs/Manuscript.Rmd
Changed formatting to submit to BioInv

---
header-includes:
  - \usepackage{natbib}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \rhead{Lionfish allometric growth}
  - \cfoot{\thepage}
title: \LARGE{Allometric growth of invasive lionfish (\textit{Pterois spp}) varies by region}
abstract:
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: no
    # template: peerj_template.tex
    toc: no
bibliography: references.bib
csl: elsevier-harvard.csl
fontsize: 12pt
---

Juan Carlos Villaseñor-Derbez^1^*, Sean Fitzgerald^1^

^1^ Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, California, United States, 93117

---

\begin{abstract}
Lionfish (\textit{Pterois volitans / miles}) are an invasive species in the Western Atlantic and the Caribbean. Improving management of invasive lionfish populations requires accurate total biomass estimates, which depend on accurate estimates of allometric growth. Sedentary species like lionfish often exhibit high levels of spatial variation in life history characterstics. We review 17 published length-weight relationships for lionfish taken throughout their invasive range and found substantial regional differences in allometric growth parameters. The spatial pattern we observed is consistent with findings from other studies focusing on genetics or age-at-length. We show that the use of \textit{ex situ} parameters can result in up to a threefold overestimation of total weight, but using parameters from nearby regions reduces this error. These findings can have major implications for management in terms of predicting effects on local ecosystems, evaluating the effectiveness of removal programs, or estimating biomass available for harvest.
\end{abstract}

---

**Keywords:** Lionfish, invasive species, length-weight, allometric, regional variations

*jvillasenor\@bren.ucsb.edu

@@

We also collected data from 10 sampling sites along the central Mexican Caribbean coast in 2010 (Table S1). Sampling locations included wall and carpet reefs at depths between 5.7 m and 38.1 m. All observed lionfish (n = 109) were collected using hand nets and numbered collection bottles. The use of hand nets prevented any weight loss due to bleeding and allowed better representation of small sizes by eliminating gear selectivity. Organisms were euthanized via pithing and Total Length (TL; mm) and Total Weight (TW; g) were recorded.

@@

@@

@@

fig_bio_ratio <- Predicted_biomass %>% 
  mutate(Study = paste0(Study, " (", Gender_site, ")")) %>%
  transform(Study = reorder(Study, -Biomass_ratio),
            Study_id = reorder(Study_id, -Biomass_ratio)) %>%
  rename(Sex = Gender) %>% 
  ggplot(aes(x = Study, y = Biomass_ratio, group = Study_id)) +
  geom_violin(aes(fill = Sex), alpha = 0.8) +
  stat_summary(geom = "point", fun.y = "mean", size = 1, color = "blue") +
  stat_summary(geom = "point", fun.y = "median", size = 1, color = "red") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(y = "Weight ratio (Predicted/Observed)", x = "") +
  coord_flip() +
  theme_cowplot() +
  theme(text = element_text(family = "Times")) +
  scale_fill_brewer(palette = "Greys") +
  theme(text = element_text(family = "Times"),
        legend.position = "top",
        legend.text = element_text(size = 10),
        legend.background = element_rect(fill = "transparent"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12)) +
  geom_text(aes(y = 3.7, label = groups), family = "Times", fontface = "plain") +
  scale_y_continuous(limits = c(NA, 3.7))
```

@@

The results presented here have major implications for management. For example, \citet{edwards_2014} simulated a lionfish culling program under two scenarios, one using length-at-age and length-to-weight parameters from North Carolina and one using parameters from Little Cayman. Their results show that using different parameters caused up to a four-year difference in the time required for the simulated lionfish population to recover to 90% of its initial biomass after removals ceased. Here, we show that using one set of length-weight parameters versus another for a given length can result in more than a threefold overestimation of TW. These spatially-driven differences become especially important when allocating resources for lionfish removal programs, incentivizing lionfish fisheries as a source of alternative livelihoods, or estimating ecosystem impacts. Research efforts focused on invasive lionfish populations need to use parameters calculated for their region to the extent possible, or at least use reasonable sets of different parameters that provide upper and lower bounds in their results.

@@

\bibliography{references}

# Figures and Tables

```{r}
review %>% 
  select(Region, Gender2, N, a, b, c, r2, Study) %>% 
  filter(!Study == "Fishbase") %>% 
  arrange(Region) %>% 
  mutate(r2 = ifelse(is.na(r2), "-", r2),
         N = ifelse(is.na(N), "-", N),
         a = formatC((a*1e5), digits = 2)) %>% 
  knitr::kable(row.names = F,
               digits =  c(0, 0, 0, 0, 2, 2),
               booktabs = T,
               # longtable = T,
               col.names = c("Region", "Sex", "n", "a", "b", "c", "Fit", "Reference"),
               caption = "\\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter for Eq. 1 (presented in $\\times 10^{-5}$), c = y-intercept for Eq. 3, b = exponent or slope for Eq. 1 or Eq. 3, respectively. The Fit column contains the reported $R^2$ of the model fit.", escape = T)
```

```{r, fig.height = 5, fig.width = 5.5, fig.cap = "\\label{fig:bio_ratio}Violin plot of predicted-to-observed weight ratios for 18 pairs of allometric parameters. Red and blue circles indicate median and mean values, respectively. Like letters indicate values that do not differ significantly."}
fig_bio_ratio
```

>>>>

---
title: Spatial variation in allometric growth of invasive lionfish has management implications
titlerunning: Lionfish allometric growth
authorrunning: Villaseñor-Derbez & Fitzgerald
authors: 
- name: Juan Carlos Villaseñor-Derbez
  address: "Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, California, United States, 93117 | ORCID: 0000-0003-1245-589X"
  email: jvillasenor@bren.ucsb.edu
- name: Sean Fitzgerald
  address: Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, California, United States, 93117

keywords:
  - Lionfish
  - invasive species
  - length-weight
  - allometric growht
  - regional variations

abstract: |
  Lionfish (\textit{Pterois volitans / miles}) are an invasive species in the Western Atlantic and the Caribbean. Improving management of invasive lionfish populations requires accurate total biomass estimates, which depend on accurate estimates of allometric growth. Sedentary species like lionfish often exhibit high levels of spatial variation in life history characterstics. We review 17 published length-weight relationships for lionfish taken throughout their invasive range and found substantial regional differences in allometric growth parameters. The spatial pattern we observed is consistent with findings from other studies focusing on genetics or age-at-length. We show that the use of \textit{ex situ} parameters can result in up to a threefold under- or overestimation of total weight, but using parameters from nearby regions reduces this error. These findings can have major implications for management in terms of predicting effects on local ecosystems, evaluating the effectiveness of removal programs, or estimating biomass available for harvest.

bibliography: references.bib
output: rticles::springer_article
header-includes:
   - \journalname{Biological Invasions}
   - \usepackage{natbib}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   # - \restylefloat{table}
---

@@

We also collected data from 10 sampling sites along the central Mexican Caribbean coast in 2010 (Supplementary Table 1). Sampling locations included wall and carpet reefs at depths between 5.7 m and 38.1 m. All observed lionfish (n = 109) were collected using hand nets and numbered collection bottles. The use of hand nets prevented any weight loss due to bleeding and allowed better representation of small sizes by eliminating gear selectivity. Organisms were euthanized via pithing and Total Length (TL; mm) and Total Weight (TW; g) were recorded.

@@

<!-- \clearpage -->

@@

```{r}
review %>% 
  select(Region, Gender2, N, a, b, c, r2, Study) %>% 
  filter(!Study == "Fishbase") %>% 
  arrange(Region) %>% 
  mutate(r2 = ifelse(is.na(r2), "-", r2),
         N = ifelse(is.na(N), "-", N),
         a = formatC((a*1e5), digits = 2)) %>% 
  knitr::kable(row.names = F,
               digits =  c(0, 0, 0, 0, 2, 2),
               # longtable = T,
               col.names = c("Region", "Sex", "n", "a", "b", "c", "Fit", "Reference"),
               caption = "\\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter for Eq. 1 (presented in $\\times 10^{-5}$), c = y-intercept for Eq. 3, b = exponent or slope for Eq. 1 or Eq. 3, respectively. The Fit column contains the reported $R^2$ of the model fit.", escape = T)
```

@@

fig_bio_ratio <- Predicted_biomass %>% 
  mutate(Study = paste0(Study, " (", Gender_site, ")")) %>%
  transform(Study = reorder(Study, -Biomass_ratio),
            Study_id = reorder(Study_id, -Biomass_ratio)) %>%
  rename(Sex = Gender) %>% 
  ggplot(aes(x = Study, y = Biomass_ratio, group = Study_id)) +
  geom_violin(aes(fill = Sex), alpha = 0.8) +
  stat_summary(geom = "point", fun.y = "median", size = 2, color = "black", fill = "steelblue", shape = 21) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(y = "Weight ratio (Predicted/Observed)", x = "") +
  coord_flip() +
  theme_cowplot() +
  theme(text = element_text(family = "Times")) +
  scale_fill_brewer(palette = "Greys") +
  theme(text = element_text(family = "Times"),
        legend.position = "top",
        legend.text = element_text(size = 10),
        legend.background = element_rect(fill = "transparent"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12)) +
  geom_text(aes(y = 3.7, label = groups), family = "Times", fontface = "plain") +
  scale_y_continuous(limits = c(NA, 3.7)) +
  guides(fill = guide_legend(title.position = "top"))
```

```{r, fig.height = 5.2, fig.width = 5.5, fig.cap = "\\label{fig:bio_ratio}Violin plot of predicted-to-observed weight ratios for 18 pairs of allometric parameters. Blue circles indicate median values and Like letters indicate values that do not differ significantly."}
fig_bio_ratio
```

\clearpage

@@

The results presented here have major implications for management. For example, \citet{edwards_2014} simulated a lionfish culling program under two scenarios, one using length-at-age and length-to-weight parameters from North Carolina and one using parameters from Little Cayman. Their results show that using different parameters caused up to a four-year difference in the time required for the simulated lionfish population to recover to 90% of its initial biomass after removals ceased. Here, we show that using one set of length-weight parameters versus another for a given length can result in more than a threefold under- or overestimation of total weight. These spatially-driven differences become especially important when allocating resources for lionfish removal programs, incentivizing lionfish fisheries as a source of alternative livelihoods, or estimating ecosystem impacts. Research efforts focused on invasive lionfish populations need to use parameters calculated for their region to the extent possible, or at least use reasonable sets of different parameters that provide upper and lower bounds in their results.

@@

Conflict of Interest: The authors declare that they have no conflict of interest.

<!-- \bibliography{references} -->

### 2018-07-13 docs/Manuscript.tex
Changed formatting to submit to BioInv

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{\LARGE{Allometric growth of invasive lionfish (\textit{Pterois spp}) varies by region}}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{natbib}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\usepackage{float}
\floatplacement{figure}{H}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\rhead{Lionfish allometric growth}
\cfoot{\thepage}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\maketitle

Juan Carlos Villaseñor-Derbez\textsuperscript{1}*, Sean
Fitzgerald\textsuperscript{1}

\textsuperscript{1} Bren School of Environmental Science and Management,
University of California Santa Barbara, Santa Barbara, California,
United States, 93117

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\begin{abstract}
Lionfish (\textit{Pterois volitans / miles}) are an invasive species in the Western Atlantic and the Caribbean. Improving management of invasive lionfish populations requires accurate total biomass estimates, which depend on accurate estimates of allometric growth. Sedentary species like lionfish often exhibit high levels of spatial variation in life history characterstics. We review 17 published length-weight relationships for lionfish taken throughout their invasive range and found substantial regional differences in allometric growth parameters. The spatial pattern we observed is consistent with findings from other studies focusing on genetics or age-at-length. We show that the use of \textit{ex situ} parameters can result in up to a threefold overestimation of total weight, but using parameters from nearby regions reduces this error. These findings can have major implications for management in terms of predicting effects on local ecosystems, evaluating the effectiveness of removal programs, or estimating biomass available for harvest.
\end{abstract}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{Keywords:} Lionfish, invasive species, length-weight,
allometric, regional variations

*jvillasenor@bren.ucsb.edu

@@

We also collected data from 10 sampling sites along the central Mexican
Caribbean coast in 2010 (Table S1). Sampling locations included wall and
carpet reefs at depths between 5.7 m and 38.1 m. All observed lionfish
(n = 109) were collected using hand nets and numbered collection
bottles. The use of hand nets prevented any weight loss due to bleeding
and allowed better representation of small sizes by eliminating gear
selectivity. Organisms were euthanized via pithing and Total Length (TL;
mm) and Total Weight (TW; g) were recorded.

@@

@@

@@

The results presented here have major implications for management. For
example, \citet{edwards_2014} simulated a lionfish culling program under
two scenarios, one using length-at-age and length-to-weight parameters
from North Carolina and one using parameters from Little Cayman. Their
results show that using different parameters caused up to a four-year
difference in the time required for the simulated lionfish population to
recover to 90\% of its initial biomass after removals ceased. Here, we
show that using one set of length-weight parameters versus another for a
given length can result in more than a threefold overestimation of TW.
These spatially-driven differences become especially important when
allocating resources for lionfish removal programs, incentivizing
lionfish fisheries as a source of alternative livelihoods, or estimating
ecosystem impacts. Research efforts focused on invasive lionfish
populations need to use parameters calculated for their region to the
extent possible, or at least use reasonable sets of different parameters
that provide upper and lower bounds in their results.

@@

\bibliography{references}

\section{Figures and Tables}\label{figures-and-tables}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{\label{fig:unnamed-chunk-8}\label{fig:map}Locations where
allometric growth parameters of lionfish (\emph{Pterois spp}) have been
reported. Circle sizes indicate sample size from each study, colors
indicate the \(b\) coefficient from Eq. \ref{eq:allometric}.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-9-1.pdf}
\caption{\label{fig:unnamed-chunk-9}\label{fig:l-w-carib}Length-weight
relationship for 109 lionfish sampled in the central Mexican Caribbean.
Points indicate samples, dashed black line indicates curve of best fit,
marginal plots represent the density distribution of each variable.}
\end{figure}

\caption{\label{tab:unnamed-chunk-11}\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter for Eq. 1 (presented in $\times 10^{-5}$), c = y-intercept for Eq. 3, b = exponent or slope for Eq. 1 or Eq. 3, respectively. The Fit column contains the reported $R^2$ of the model fit.}
\centering
\begin{tabular}[t]{llllrrll}
\toprule
Region & Sex & n & a & b & c & Fit & Reference\\
\midrule
Caribbean & B & 458 & 3.6 & 2.81 & -4.44 & - & Sandel et al., 2015\\
Caribbean & B & 419 & 2.8 & 2.85 & -4.56 & 0.8715 & Chin et al., 2016\\
Caribbean & B & 1450 & 2.3 & 2.89 & -4.64 & 0.96 & de Leon et al., 2013\\
Caribbean & B & 1887 & 0.3 & 3.24 & -5.52 & 0.97 & Edwards et al., 2014\\
Caribbean & B & - & 0.25 & 3.29 & -5.60 & - & Darling et al., 2011\\
\addlinespace
Caribbean & B & 2143 & 0.52 & 3.18 & -5.28 & 0.9907 & Sabido-Itza et al., 2016\\
Caribbean & B & 227 & 0.8 & 3.11 & -5.10 & 0.958 & Toledo-Hernández et al., 2014\\
Caribbean & B & 449 & 0.23 & 3.25 & -5.64 & 0.97 & Sabido-Itza et al., 2016b\\
Caribbean & B & 368 & 0.32 & 3.19 & -5.50 & 0.98 & Sabido-Itza et al., 2016b\\
Caribbean & B & 109 & 0.32 & 3.23 & -5.49 & 0.9766 & This study\\
\addlinespace
GoM & B & 934 & 0.21 & 3.34 & -5.68 & 0.98 & Dahl \& Patterson, 2014\\
GoM & B & 472 & 0.29 & 3.30 & -5.54 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & F & 67 & 0.12 & 3.47 & -5.93 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & M & 59 & 0.42 & 3.23 & -5.38 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & B & 582 & 0.14 & 3.43 & -5.86 & 0.99 & Fogg et al., 2013\\
\addlinespace
GoM & M & 119 & 0.27 & 3.31 & -5.57 & 0.97 & Fogg et al., 2013\\
GoM & F & 115 & 0.68 & 3.14 & -5.17 & 0.94 & Fogg et al., 2013\\
North Atlantic & B & 774 & 2.9 & 2.89 & -4.54 & - & Barbour et al., 2011\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-12-1.pdf}
\caption{\label{fig:unnamed-chunk-12}\label{fig:all_allo}Length-weight
relationships (n = 18) for 12 studies and this study. Colors indicate
studies from which the parameters were extracted. Dotted, dashed and
solid lines show models for males, females, and combined sexes,
respectively. The dashed black line represents the relationship
estimated in this study.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-13-1.pdf}
\caption{\label{fig:unnamed-chunk-13}\label{fig:bio_ratio}Violin plot of
predicted-to-observed weight ratios for 18 pairs of allometric
parameters. Red and blue circles indicate median and mean values,
respectively. Like letters indicate values that do not differ
significantly.}
\end{figure}

>>>>

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

\journalname{Biological Invasions}
\usepackage{natbib}
\usepackage{float}
\floatplacement{figure}{H}

\begin{document}

\title{Spatial variation in allometric growth of invasive lionfish has
management implications }

    \titlerunning{Lionfish allometric growth}

\author{  Juan Carlos Villaseñor-Derbez \and  Sean Fitzgerald \and  }

    \authorrunning{ Villaseñor-Derbez \& Fitzgerald }

\institute{
        Juan Carlos Villaseñor-Derbez \at
     Bren School of Environmental Science and Management, University of
 California Santa Barbara, Santa Barbara, California, United States,
 93117 \textbar{} ORCID: 0000-0003-1245-589X \\
     \email{\href{mailto:jvillasenor@bren.ucsb.edu}{\nolinkurl{jvillasenor@bren.ucsb.edu}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Sean Fitzgerald \at
     Bren School of Environmental Science and Management, University of
 California Santa Barbara, Santa Barbara, California, United States,
 93117 \\
     %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Lionfish (\textit{Pterois volitans / miles}) are an invasive species in
the Western Atlantic and the Caribbean. Improving management of invasive
lionfish populations requires accurate total biomass estimates, which
depend on accurate estimates of allometric growth. Sedentary species
like lionfish often exhibit high levels of spatial variation in life
history characterstics. We review 17 published length-weight
relationships for lionfish taken throughout their invasive range and
found substantial regional differences in allometric growth parameters.
The spatial pattern we observed is consistent with findings from other
studies focusing on genetics or age-at-length. We show that the use of
\textit{ex situ} parameters can result in up to a threefold under- or
overestimation of total weight, but using parameters from nearby regions
reduces this error. These findings can have major implications for
management in terms of predicting effects on local ecosystems,
evaluating the effectiveness of removal programs, or estimating biomass
available for harvest.
\\
\keywords{
        Lionfish \and
        invasive species \and
        length-weight \and
        allometric growht \and
        regional variations \and
    }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

@@

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:map}Locations where allometric growth parameters of
lionfish (\emph{Pterois spp}) have been reported. Circle sizes indicate
sample size from each study, colors indicate the \(b\) coefficient from
Eq. \ref{eq:allometric}.}
\end{figure}

We also collected data from 10 sampling sites along the central Mexican
Caribbean coast in 2010 (Supplementary Table 1). Sampling locations
included wall and carpet reefs at depths between 5.7 m and 38.1 m. All
observed lionfish (n = 109) were collected using hand nets and numbered
collection bottles. The use of hand nets prevented any weight loss due
to bleeding and allowed better representation of small sizes by
eliminating gear selectivity. Organisms were euthanized via pithing and
Total Length (TL; mm) and Total Weight (TW; g) were recorded.

@@

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{\label{fig:l-w-carib}Length-weight relationship for 109
lionfish sampled in the central Mexican Caribbean. Points indicate
samples, dashed black line indicates curve of best fit, marginal plots
represent the density distribution of each variable.}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-9}\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter for Eq. 1 (presented in $\times 10^{-5}$), c = y-intercept for Eq. 3, b = exponent or slope for Eq. 1 or Eq. 3, respectively. The Fit column contains the reported $R^2$ of the model fit.}
\centering
\begin{tabular}[t]{l|l|l|l|r|r|l|l}
\hline
Region & Sex & n & a & b & c & Fit & Reference\\
\hline
Caribbean & B & 458 & 3.6 & 2.81 & -4.44 & - & Sandel et al., 2015\\
\hline
Caribbean & B & 419 & 2.8 & 2.85 & -4.56 & 0.8715 & Chin et al., 2016\\
\hline
Caribbean & B & 1450 & 2.3 & 2.89 & -4.64 & 0.96 & de Leon et al., 2013\\
\hline
Caribbean & B & 1887 & 0.3 & 3.24 & -5.52 & 0.97 & Edwards et al., 2014\\
\hline
Caribbean & B & - & 0.25 & 3.29 & -5.60 & - & Darling et al., 2011\\
\hline
Caribbean & B & 2143 & 0.52 & 3.18 & -5.28 & 0.9907 & Sabido-Itza et al., 2016\\
\hline
Caribbean & B & 227 & 0.8 & 3.11 & -5.10 & 0.958 & Toledo-Hernández et al., 2014\\
\hline
Caribbean & B & 449 & 0.23 & 3.25 & -5.64 & 0.97 & Sabido-Itza et al., 2016b\\
\hline
Caribbean & B & 368 & 0.32 & 3.19 & -5.50 & 0.98 & Sabido-Itza et al., 2016b\\
\hline
Caribbean & B & 109 & 0.32 & 3.23 & -5.49 & 0.9766 & This study\\
\hline
GoM & B & 934 & 0.21 & 3.34 & -5.68 & 0.98 & Dahl \& Patterson, 2014\\
\hline
GoM & B & 472 & 0.29 & 3.30 & -5.54 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
\hline
GoM & F & 67 & 0.12 & 3.47 & -5.93 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
\hline
GoM & M & 59 & 0.42 & 3.23 & -5.38 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
\hline
GoM & B & 582 & 0.14 & 3.43 & -5.86 & 0.99 & Fogg et al., 2013\\
\hline
GoM & M & 119 & 0.27 & 3.31 & -5.57 & 0.97 & Fogg et al., 2013\\
\hline
GoM & F & 115 & 0.68 & 3.14 & -5.17 & 0.94 & Fogg et al., 2013\\
\hline
North Atlantic & B & 774 & 2.9 & 2.89 & -4.54 & - & Barbour et al., 2011\\
\hline
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-11-1.pdf}
\caption{\label{fig:all_allo}Length-weight relationships (n = 18) for 12
studies and this study. Colors indicate studies from which the
parameters were extracted. Dotted, dashed and solid lines show models
for males, females, and combined sexes, respectively. The dashed black
line represents the relationship estimated in this study.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-13-1.pdf}
\caption{\label{fig:bio_ratio}Violin plot of predicted-to-observed
weight ratios for 18 pairs of allometric parameters. Blue circles
indicate median values and Like letters indicate values that do not
differ significantly.}
\end{figure}

\clearpage

@@

The results presented here have major implications for management. For
example, \citet{edwards_2014} simulated a lionfish culling program under
two scenarios, one using length-at-age and length-to-weight parameters
from North Carolina and one using parameters from Little Cayman. Their
results show that using different parameters caused up to a four-year
difference in the time required for the simulated lionfish population to
recover to 90\% of its initial biomass after removals ceased. Here, we
show that using one set of length-weight parameters versus another for a
given length can result in more than a threefold under- or
overestimation of total weight. These spatially-driven differences
become especially important when allocating resources for lionfish
removal programs, incentivizing lionfish fisheries as a source of
alternative livelihoods, or estimating ecosystem impacts. Research
efforts focused on invasive lionfish populations need to use parameters
calculated for their region to the extent possible, or at least use
reasonable sets of different parameters that provide upper and lower
bounds in their results.

@@

Conflict of Interest: The authors declare that they have no conflict of
interest.

\bibliographystyle{spbasic}
\bibliography{references.bib}

### 2018-07-13 docs/supplementary/S1Table.Rmd
Changed formatting to submit to BioInv

---
title: "Supplementary Table S1"
author: "Juan Carlos Villaseñor-Derbez"
output: pdf_document
---

@@

```{r}
fish <- read_csv("./lionfish_data.csv", col_types = cols()) %>% 
  group_by(Id, Site, Location, Latitude, Longitude, Depth_m) %>% 
  summarize(Length = 10*mean(TotalLength_cm),
            Weight = mean(TotalWeigth_gr),
            Stomach = mean(StomachLength_cm)) %>% 
  ungroup()
```

@@

fish %>% 
  group_by(Location) %>% 
  summarize(Latitude = formatC(mean(Latitude), digits = 3, format = "f"),
            Longitude = formatC(mean(Longitude), digits = 3, format = "f"),
            minD = formatC(min(Depth_m), digits = 1, format = "f"),
            maxD = formatC(max(Depth_m), digits = 1, format = "f"),
            meanD = formatC(mean(Depth_m), digits = 1, format = "f"),
            n = n()) %>% 
  rbind(total) %>% 
  knitr::kable(row.names = F,
               col.names = c("Location", "Lat.", "Long.", "Min. Depth", "Max. Depth", "Mean Depth", "n"))
```

>>>>

---
title: Spatial variation in allometric growth of invasive lionfish has management implications
subtitle: "Supplementary Table 1"
author: "Villaseñor-Derbez & Fitzgerald"
output: bookdown::pdf_document2
fontsize: 12pt
---

@@

```{r}
fish <- read_csv(here::here("data", "lionfish_data.csv"), col_types = cols()) %>% 
  group_by(Id, Site, Location, Latitude, Longitude, Depth_m) %>% 
  summarize(Length = 10*mean(TotalLength_cm),
            Weight = mean(TotalWeigth_gr),
            Stomach = mean(StomachLength_cm)) %>% 
  ungroup()
```

@@

fish %>% 
  group_by(Location) %>% 
  summarize(Latitude = formatC(mean(Latitude), digits = 3, format = "f"),
            Longitude = formatC(mean(Longitude), digits = 3, format = "f"),
            minD = formatC(min(Depth_m), digits = 1, format = "f"),
            maxD = formatC(max(Depth_m), digits = 1, format = "f"),
            meanD = formatC(mean(Depth_m), digits = 1, format = "f"),
            n = n()) %>% 
  rbind(total) %>% 
  knitr::kable(row.names = F,
               booktabs = T,
               col.names = c("Location", "Lat.", "Long.", "Min. Depth", "Max. Depth", "Mean Depth", "n"))
```
